---
title: "Analysing the Russia Report"
author: "Andi Fugard (almost@gmail.com, @[inductivestep](https://twitter.com/InductiveStep))"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    self_contained: no
    #toc: yes
    #toc_float: yes
---

The report is available on the [ISC website](http://isc.independent.gov.uk/news-archive/21july2020) and [in this github repository](20200721_HC632_CCS001_CCS1019402408-001_ISC_Russia_Report_Web_Accessible.pdf).

First, load it into R.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(pdftools)
library(stringr)
rr_text <- pdf_text("20200721_HC632_CCS001_CCS1019402408-001_ISC_Russia_Report_Web_Accessible.pdf")
```
The `pdf_text` command returns a character vector with one element per PDF page. The main report text is on pages 8-42.


I'd like to know where the juicy pages are - one way to index them is by number of reduction marks, ***.

First try for one page to get the hang of it. Page 10 has one.

```{r}
str_count(rr_text[10], pattern = "\\*\\*\\*")
```

It worked - hurrah. Try for page 8 which has none.

```{r}
str_count(rr_text[8], pattern = "\\*\\*\\*")
```

And page 12, which has 8:

```{r}
str_count(rr_text[12], pattern = "\\*\\*\\*")
```

All good. `str_count` takes a vector, so we can do the whole report in one go:

```{r}
redact_n <- str_count(rr_text, pattern = "\\*\\*\\*")
```

Let's stick these in a tibble, alongside the (pdf, as opposed to printed) page number and original page text.

```{r}
rr_tib <- tibble(page = 1:length(rr_text),
                 text = rr_text,
                 redactions = redact_n)
```

Now a histogram of those redactions:

```{r message=FALSE}
ggplot(rr_tib, aes(x=redactions)) +
  geom_histogram(binwidth = 1, fill = "firebrick", color = "white") +
  labs(x = "Number of redactions",
       y = "Count",
       title = "Histogram of redactions on each page of the Russia Report") +
  theme_classic()
```

What page has the maximum number of redactions?

```{r}
rr_tib %>%
  filter(redactions == max(rr_tib$redactions)) %>%
  select(page, redactions)
```

## Ta da - the page with most redactions (p. 31 in the printed report)

```{r results='asis'}
cat(gsub(pattern = "\\*\\*\\*",
         replace = "\\*\\*\\*[redacted]\\*\\*\\*",
         x = rr_tib$text[38]))
```





